---
title: "Inference for camera trap data using 'camtrap'."
author: "James Henderson"
date: "2/20/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

The $\tt{R}$ package $\tt{camtrap}$ is built on and extends the capabilities of
the $\tt{overlap}$ package for analyzing time-of-day data as generated by camera 
traps. The primary extensions are:
* modified functions allowing attention to be focused on a specific time windows,
* the addition of estimates of the area between densities rather than only the
overlap,
* the addition of "permutation" tests for testing hypotheses. 
The following sections are meant to illusrtate the use of the added functionality.

## Quick Start
In this section we briefly describe how to get up and running with this package.
This is a develoment version of the package available on github. For easy installation,
first install and load the $\tt{devtools}$ package.

```{r devtools}
#install.packages('devtools')
library(devtools)
```

You can then install $\tt{camtrap}$ directly from github.

```{r camtrap}
install_git('github.com/jbhender/camtrap/)
```
## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
